
##digitrecognizer

Neeraj
December 27, 2015

```{r}
#install.packages("readr")
library(readr)
## Warning: package 'readr' was built under R version 3.2.3
train <- read_csv("train.csv")
test <- read_csv("test.csv")
head(train[1:10])
```

```
##   label pixel0 pixel1 pixel2 pixel3 pixel4 pixel5 pixel6 pixel7 pixel8
## 1     1      0      0      0      0      0      0      0      0      0
## 2     0      0      0      0      0      0      0      0      0      0
## 3     1      0      0      0      0      0      0      0      0      0
## 4     4      0      0      0      0      0      0      0      0      0
## 5     0      0      0      0      0      0      0      0      0      0
## 6     0      0      0      0      0      0      0      0      0      0
# Create a 28*28 matrix with pixel color values
m = matrix(unlist(train[10,-1]),nrow = 28,byrow = T)
# Plot that matrix
image(m,col=grey.colors(255))

```


This image needs to be rotated to the right. I will rotate the matrix and plot a bunch of images.
rotate <- function(x) t(apply(x, 2, rev)) # reverses (rotates the matrix)

```{r}

# Plot a bunch of images
par(mfrow=c(2,3))
lapply(1:6, 
    function(x) image(
                    rotate(matrix(unlist(train[x,-1]),nrow = 28,byrow = T)),
                    col=grey.colors(255),
                    xlab=train[x,1]
                )
)

```


```

## [[1]]
## NULL
## 
## [[2]]
## NULL
## 
## [[3]]
## NULL
## 
## [[4]]
## NULL
## 
## [[5]]
## NULL
## 
## [[6]]
## NULL

```

```{r}

par(mfrow=c(1,1)) # set plot options back to default
library(h2o)
## Loading required package: statmod
## 
## ----------------------------------------------------------------------
## 
## Your next step is to start H2O and get a connection object (named
## 'localH2O', for example):
##     > localH2O = h2o.init()
## 
## For H2O package documentation, ask for help:
##     > ??h2o
## 
## After starting H2O, you can use the Web UI at http://localhost:54321
## For more information visit http://docs.h2o.ai
## 
## ----------------------------------------------------------------------
## start a local h2o cluster
localH2O = h2o.init(#max_mem_size = '6g', # use 6GB of RAM of *GB available
                    nthreads = -1) # use all CPUs (8 on my personal computer :3)
## 
## H2O is not running yet, starting it now...
## Warning in .h2o.startJar(nthreads = nthreads, max_memory = max_mem_size, : You have a 32-bit version of Java. H2O works best with 64-bit Java.
## Please download the latest Java SE JDK 7 from the following URL:
## http://www.oracle.com/technetwork/java/javase/downloads/jdk7-downloads-1880260.html
## 
## Note:  In case of errors look at the following log files:
##     C:\Users\NJ_NEE~1\AppData\Local\Temp\RtmpkFuuvK/h2o_Nj_neeraj_started_from_r.out
##     C:\Users\NJ_NEE~1\AppData\Local\Temp\RtmpkFuuvK/h2o_Nj_neeraj_started_from_r.err
## 
## 
## .Successfully connected to http://127.0.0.1:54321/ 
## 
## R is connected to the H2O cluster: 
##     H2O cluster uptime:         2 seconds 912 milliseconds 
##     H2O cluster version:        3.2.0.3 
##     H2O cluster name:           H2O_started_from_R_Nj_neeraj_fut736 
##     H2O cluster total nodes:    1 
##     H2O cluster total memory:   0.97 GB 
##     H2O cluster total cores:    4 
##     H2O cluster allowed cores:  4 
##     H2O cluster healthy:        TRUE

```


Now, I just convert the train and test sets into the h2o format and set up the model. The h2o.deeplearning() function has lots of configurable arguments. For demonstration I went with a two layer neural network with 100 nodes and 0.5 dropout ratio in each. While I could specify a learning rate I decided not to because it is adaptive by default which should result in an improved accuracy score. This blog post explores the deeplearning using h2o and its many configurable options. I used it for reference. So let’s train the model.

```{r}

## MNIST data as H2O
train[,1] = as.factor(train[,1]) # convert digit labels to factor for classification
train_h2o = as.h2o(train)
## 
  |                                                                       
  |                                                                 |   0%
  |                                                                       
  |=================================================================| 100%
test_h2o = as.h2o(test)
## 
  |                                                                       
  |                                                                 |   0%
  |                                                                       
  |=================================================================| 100%
## set timer
s <- proc.time()

## train model
model =
  h2o.deeplearning(x = 2:785,  # column numbers for predictors
                   y = 1,   # column number for label
                   training_frame = train_h2o, # data in H2O format
                   activation = "RectifierWithDropout", # algorithm
                   input_dropout_ratio = 0.2, # % of inputs dropout
                   hidden_dropout_ratios = c(0.5,0.5), # % for nodes dropout
                   balance_classes = TRUE, 
                   hidden = c(100,100), # two layers of 100 nodes
                   momentum_stable = 0.99,
                   nesterov_accelerated_gradient = T, # use it for speed
                   epochs = 15) # no. of epochs
## 
  |                                                                       
  |                                                                 |   0%
## Warning in .h2o.validateModelParameters(conn, algo, param_values, h2oRestApiVersion): Dropping constant columns: pixel0, pixel1, pixel2, pixel3, pixel4, pixel5, pixel6, pixel7, pixel8, pixel9, pixel10, pixel11, pixel16, pixel17, pixel18, pixel19, pixel20, pixel21, pixel22, pixel23, pixel24, pixel25, pixel26, pixel27, pixel28, pixel29, pixel30, pixel31, pixel52, pixel53, pixel54, pixel55, pixel56, pixel57, pixel82, pixel83, pixel84, pixel85, pixel111, pixel112, pixel139, pixel140, pixel141, pixel168, pixel196, pixel392, pixel420, pixel421, pixel448, pixel476, pixel532, pixel560, pixel644, pixel645, pixel671, pixel672, pixel673, pixel699, pixel700, pixel701, pixel727, pixel728, pixel729, pixel730, pixel731, pixel754, pixel755, pixel756, pixel757, pixel758, pixel759, pixel760, pixel780, pixel781, pixel782, pixel783.
## 
  |                                                                       
  |===                                                              |   5%
  |                                                                       
  |======                                                           |  10%
  |                                                                       
  |==========                                                       |  15%
  |                                                                       
  |=============                                                    |  20%
  |                                                                       
  |================                                                 |  25%
  |                                                                       
  |===================                                              |  30%
  |                                                                       
  |======================                                           |  35%
  |                                                                       
  |==========================                                       |  39%
  |                                                                       
  |=============================                                    |  44%
  |                                                                       
  |================================                                 |  49%
  |                                                                       
  |===================================                              |  54%
  |                                                                       
  |=======================================                          |  59%
  |                                                                       
  |==========================================                       |  64%
  |                                                                       
  |=============================================                    |  69%
  |                                                                       
  |================================================                 |  74%
  |                                                                       
  |===================================================              |  79%
  |                                                                       
  |=======================================================          |  84%
  |                                                                       
  |==========================================================       |  89%
  |                                                                       
  |=============================================================    |  94%
  |                                                                       
  |================================================================ |  99%
  |                                                                       
  |=================================================================| 100%
## print confusion matrix
h2o.confusionMatrix(model)
## Confusion Matrix - (vertical: actual; across: predicted): vertical: actual; across: predicted
##           0   1    2    3    4   5    6    7    8   9  Error
## 0       987   0    2    2    1   1    3    1    3   0 0.0130
## 1         0 956    7    3    1   0    1    6    3   0 0.0215
## 2         0   2  979    5    3   1    3   31    4   1 0.0486
## 3         1   0   13  974    0   6    0   19    2   3 0.0432
## 4         0   1   11    0  997   1    8    7    0  14 0.0404
## 5         5   0    6   24    2 962   12    6    4   1 0.0587
## 6         5   1   10    0    1   2  986   18    0   0 0.0362
## 7         1   1    8    0    2   1    1  953    0   1 0.0155
## 8         2   5   10    4    2   3    3    6 1011   5 0.0381
## 9         2   2    1    8   15   5    0   39    3 898 0.0771
## Totals 1003 968 1047 1020 1024 982 1017 1086 1030 923 0.0393
##                  Rate
## 0      =   13 / 1,000
## 1      =     21 / 977
## 2      =   50 / 1,029
## 3      =   44 / 1,018
## 4      =   42 / 1,039
## 5      =   60 / 1,022
## 6      =   37 / 1,023
## 7      =     15 / 968
## 8      =   40 / 1,051
## 9      =     75 / 973
## Totals = 397 / 10,100
```



After training the model we can look at the confusion matrix. The total error after 15 epochs is around 0.0395784 which translates to 1 - 0.0395784 = 0.9604216 accuracy score. The training process lasted for -42.966 seconds. Good performance within a short period of time.



## print time elapsed

```{r}

s - proc.time()
##    user  system elapsed 
##   -6.22   -0.22 -359.95

```


Now, let’s predict the test data and save the results.

```{r}

## classify test set
h2o_y_test <- h2o.predict(model, test_h2o)

## convert H2O format into data frame and  save as csv
df_y_test = as.data.frame(h2o_y_test)
df_y_test = data.frame(ImageId = seq(1,length(df_y_test$predict)), Label = df_y_test$predict)
write.csv(df_y_test, file = "submission-r-h2o.csv", row.names=F)

## shut down virutal H2O cluster
h2o.shutdown(prompt = F)
## [1] TRUE
```
